\section{}





\subsection{}
\label{classification of simple modules}

Let more generally~$R$ be any ring and let~$M$ be an~{\module{$R$}}.
Recall that a submodule~$N \subseteq M$ is \emph{maximal} if it follows for every intermediate submodule~$N \subseteq P \subseteq M$ that~$P= N$ or~$P = M$.
This is equivalent to the quotient module~$M/N$ having precisely two submodules, i.e. equivalent to~$M/N$ being simple.
It follows in particular that~$R/M$ is simple for every maximal left ideal~$M \subseteq R$.

Every simple~\dash{$R$}{module}~$S$ is already of this form.
The module~$S$ is cyclic:
It holds that~$S \neq 0$ and for~$x \in S$ with~$x \neq 0$ the submodule~$\gen{x}$ is a nonzero submodule of~$S$.
Hence~$S = \gen{x}$ because~$S$ is simple.
This shows that~$S \cong R/M$ for some module~$M$.
The simplicity of~$S$ is by the above argumentation equivalent to~$M$ being maximal.

The maximal ideals in~$\Integer$ are~$p \Integer$ with~$p$ prime.
The simple~{\modules{$\Integer$}} are therefore (up to isomorphism) precisely~$\Integer/p\Integer$ with~$p$ prime.

\begin{remark}
  We have seen above that a simple module~$S$ is not only cyclic, but that every nonzero element~$x \in S$ is already a cyclic generator.
  This is an equivalent characterization of simple modules:
  A module~$S$ is simple if and only if~$S \neq 0$ and every nonzero~$x \in S$ is a cyclic generator.
  
  Indeed, if~$S$ satisfies this condition(s), then~$S$ contains no proper nonzero cyclic submodule, and hence no proper nonzero submodule.
  Since~$S \neq 0$, this means that~$S$ is simple.
\end{remark}

\begin{warning}
  Different maximal ideals~$M, M' \subseteq R$ may given isomorphic simple modules~$R/M \cong R/M'$.
  An example for this is~$R = \mat{n}{k}$ with~$n \geq 2$ and~$k$ a field.
\end{warning}





\subsection{}

If~$R$ is an integral domain that is not a field then~$R$ not semisimple:
It holds for any two nonzero ideals~$I, J \subseteq R$ that~$I \cap J \supseteq IJ \neq 0$, and hence that the sum~$I + J$ is not direct.
If~$x \in R$ is a nonzero \dash{non}{unit} then this shows that the generated ideal~$\gen{x}$ has no direct complement.

In particular~$2 \Integer \subseteq \Integer$ has no direct complement.





\subsection{}

Every semisimple~\dash{$\Integer$}{module}~$M$ is by part~\ref*{classification of simple modules} of the form
\[
  M \cong \bigoplus_{i \in I} \Integer/p_i
\]
for some primes~$p_i$.
The primes~$p_i$ are in particular \dash{square}{free}, which proves the statement.


\begin{remark}
  One has for every~$n \geq 0$ that~$\Integer/n$ is semisimple if and only if~$n$ is \dash{square}{free}.
  Indeed, if~$n = 0$ then~$\Integer/n = \Integer$ is not semisimple, and if~$n = 1$ then~$\Integer/n = 0$ is semisimple.
  For~$n \geq 2$ we have~$n = p_1^{n_1} \dotsm p_r^{n_r}$ for some pairwise different primes~$p_i$ and exponents~$n_i \geq 1$.
  Then
  \[
          \Integer/n
    \cong \Integer/p_1^{n_1} \oplus \dotsb \oplus \Integer/p_r^{n_r}
  \]
  by the chinese reminder theorem.
  This is already a decomposition into indecomposable~{\modules{$\Integer$}} by the classification of finitely generated abelian groups.
  Hence~$\Integer/n$ is semisimple if and only if every summand~$\Integer/p_i^{n_i}$ is already simple.
  We have seen above that this is the case if and only if every~$p_i^{n_i}$ is prime, i.e.\ if and only if~$n_i = 1$ for every~$i = 1, \dotsc, r$.
  
  It follows that a~{\module{$\Integer$}}~$M$ is semisimple if and only if~$M \cong \bigoplus_i \Integer/n_i$ where the~$n_i$ are \dash{square}{free} integers.
\end{remark}


